# Ch 5. The Many Variables & The Spurious Waffles {#mcelreath_ch5}

> Reasons given for multiple regression models include: (1) Statistical -control” for confounds

DR: but endogenous controls can make this problem worse

## 5.1. Spurious association

> But there’s no reason high marriage rate must cause more divorce

DR: what would this even mean ... the 'causal' part is ambiguous in this example?

> easy to imagine high marriage rate indicating high cultural valuation of marriage and therefore being associated with low divorce rate

DR: but the latter is not causal

$\alpha ∼ Normal(0, 0.2)$ -- because with standardized variables we should have an intercept near-0

DR: why not exactly 0?

> Since the outcome and the predictor are both standardized, the intercept - should end up very close to zero DR:

> So when $\beta_A = 1$, a change of 1.2 years in median age at marriage is associated with a full standard deviation change in the outcome variable  [which he thinks is absurdly large]
gg
DR: But why absurdly large, maybe the sd in the outcome is small too?

> [a model] that includes both age at marriage and marriage rate will help us. 5

DR: maybe but these are aggregates. casual thinking will be difficult.  primitives are more. like'whether an individual of a.certain age  gets married or divorced'

> causal impact

DR: redundant language. All impacts are causal :)

> DAG will tell you the consequences of intervening to change a variable. But only if the DAG is correct

> Perhaps a direct effect would arise because younger people change faster than older people and are therefore more likely to grow incompatible with a partner. Second, it has an indirect effect by influencing the marriage rate, which then influences divorce, A $\rightarrow$  M $\rightarrow$  D

> these different arrows, we need more than one statistical model.

> Model m5.1, the regression of D on A, tells us only that the total influence of age at marriage is strongly negative with divorce rate

DR: May not be *influence*... because unobserved factor could drive both

> The 'total' here means we have to account for every path from A to D. There are two such paths in this graph: $A \rightarrow  D$, a direct path, and $A \rightarrow  M \rightarrow D$, an indirect path. In general, it is possible that a variable like A has no direct effect at all on an outcome like D. It could still be associated with D entirely through the indirect path. That type of relationship is known as mediation, and we'll have another example later.



> This DAG is also consistent with the posterior distributions of models m5.1 and m5.2. Why? Because both M and D 'listen' to A

> conditional independencies  ... which variables become dis-associated when we condition on some other set of variables. D

... that every pair of variables is correlated. This is because there is a causal arrow between every pair. These arrows create correlations.

> DR: but effects could add up to zero by coincidence. (Actually, he has a case like this further down, he just fails to mention it here)

> They share a cause, and this leads them to be correlated with one another through that cause. But suppose we condition on A. All of the information in M that is relevant to predicting D is in A. So once we've conditioned on A, M tells us nothing more about D. So in the second DAG, a testable implication is that D is independent of M, conditional on A. In other words, $D \perp  M|A$.

> Here’s the code to define the second DAG and display the implied conditional independencies.

```{r}

DMA_dag2 <- dagitty('dag{ D <- A -> M }')
impliedConditionalIndependencies( DMA_dag2 )
```

> So for example once you fit a multiple regression to predict divorce using both marriage rate and age at marriage, the model addresses the questions: (1) After I already know marriage rate, what additional value is there in also knowing age at marriage? (2) After I already know age at marriage, what additional value is there in also knowing marriage rate

DR: can we differentiate 1 from 2?


### 5.1.3. Multiple regression notation

Model with

`mu <- a + bM*M + bA*A`
> Notice how bA doesn’t move, only grows a bit more uncertain, while bM is only associated with divorce when age at marriage is missing from the model. You can interpret these distributions as saying:
> Once we know median age at marriage for a State, there is little or no additional predictive power in also knowing the rate of marriage in that State.
> In that weird notation, $D \perp M|A$. This tests the implication of the second DAG from earlier.
Since the first DAG did not imply this result, it is out

DR: This seems like a very weak test and a poor strategy for making inferences about causality (more so *running* with those inferences to underly future modeling).  Even when age *is* present in the model, the compatibility intervals for the marriage rate coefficient include rather large coefficients in either direction.


### 5.1.5. Plotting multivariate posteriors.


> (1)  Predictor residual plots. These plots show the outcome against residual predictor
values. ...
(2) Posterior prediction plots. These show model-based predictions against raw data,
or otherwise display the error in prediction. They are tools for checking fit and
assessing predictions. ...
(3) Counterfactual plots. These show the implied predictions for imaginary experiments. These plots allow you to explore the causal implications of manipulating one or more variables.



**Skipped a bunch here; come back to it**

> The trick with simulating counterfactuals is to realize that when we manipulate some variable X, we break the causal influence of other variables on X.

> Now we can use sim, which you met in the previous chapter, to simulate observations from model `m5.3_A`. But this time we’ll tell it to simulate both M and D, in that order. Why in that order? Because we have to simulate the influence of A on M before we simulate the joint influence of A and M on D. The vars argument to sim tells it both which observables to simulate and in which order.

```{r}
# prep data 5.21
sim_dat <- data.frame( A=A_seq )
# simulate M and then D, using A_seq
s <- sim( m5.3_A , data=sim_dat , vars=c("M","D") )
```
```{r}
plot( sim_dat$A , colMeans(s$D) , ylim=c(-2,2) , type="l" ,
xlab="manipulated A" , ylab="counterfactual D" )
shade( apply(s$D,2,PI) , sim_dat$A )
mtext( "Total counterfactual effect of A on D" )
```
