# Survey design, item dev. {#surveydesign}

This section will cover specific 'survey' related content, as well as some content that overlaps with experiments and trials.^M[ore technical and involved discussion of overlapping content will often be deferred to the 'experiments' related sections, with placeholders and links.]


::: {.alert .alert-secondary}


Link or incorporate William Elsey's work [HERE](https://willemsleegers.github.io/how-to-science/content/methodology/survey-design/),

and [Reinstein's sloppy notes here](https://daaronr.github.io/metrics_discussion/surveys.html) focusing on probability sampling and sampling rare populations; see embeds below


:::


## (Willem Sleegers, to incorporate)

[item development](https://willemsleegers.github.io/how-to-science/content/methodology/survey-design/item-development.html)

[wayback link](https://web.archive.org/web/20220630084700/https://willemsleegers.github.io/how-to-science/content/methodology/survey-design/item-development.html)

```{r }

knitr::include_url("https://willemsleegers.github.io/how-to-science/content/methodology/survey-design/item-development.html")

```

## Reinstein misc notes to incorporate {-}

```{r }

knitr::include_url("https://daaronr.github.io/metrics_discussion/surveys.html")

```

## Suggested sections...

**'Qualitative' issues** (hopefully backed by evidence)

- How to ask good survey questions

- Avoiding pitfalls (framing, agreeability bias, comprehension, attention,  etc.)

- Validating the measure is capturing 'the thing you intend'

\

**'Quantitative issues'**

- Constructing reliable indices and scales

- Sampling issues and representativeness
   - Elsey on "Mr P"
   - Reinstein on ['sampling'](https://daaronr.github.io/metrics_discussion/surveys.html#survey-samplingintake) and ['measuring a rare population'](https://daaronr.github.io/metrics_discussion/surveys.html#jazz-case) with reference to the EA survey

\

**'Implementation'**

- Participant platforms

- Implementation platforms and issues

- Survey design tools (IT)




## Discussions to integrate


- Disputing claim that "One doesn't need to worry about framing effects" because of Morewedge et al, 2015; Khan et al, 2006; Baumer et al, 2015 [private slack conversation link](https://rethinkpriorities.slack.com/archives/G01962YABHB/p1655407493604669)
