# Survey design, item dev. {#surveydesign}

This section will cover specific 'survey' related content, as well as some content that overlaps with experiments and trials.

More technical and involved discussion of overlapping content will often be deferred to the [experiments and trials](#expt_trial) section, with placeholders and links.

@schwarcz2007


::: {.alert .alert-secondary}


Link or incorporate William Elsey's work [HERE](https://willemsleegers.github.io/how-to-science/survey-design.html#item-development),

and [Reinstein's sloppy notes](https://daaronr.github.io/metrics_discussion/surveys.html); see embeds below


:::


## (Willem Sleegers, to incorporate)

[item development](https://willemsleegers.github.io/how-to-science/survey-design.html#item-development)

```{r }

knitr::include_url("https://willemsleegers.github.io/how-to-science/survey-design.html#item-development")

```

## Reinstein misc notes to incorporate {-}

```{r }

knitr::include_url("https://daaronr.github.io/metrics_discussion/surveys.html")

```

## Suggested sections...

- How to ask good survey questions

- Avoiding pitfalls

- Constructing reliable indices and scales

- Sampling issues and representativeness
   - Elsey on "Mr P"
   - Reinstein on ['sampling'](https://daaronr.github.io/metrics_discussion/surveys.html#survey-samplingintake) and ['measuring a rare population'](https://daaronr.github.io/metrics_discussion/surveys.html#jazz-case) with reference to the EA survey


- Implementation platforms and issues

- Survey design tools (IT)


## Discussions to integrate


- Disputing claim that "One doesn't need to worry about framing effects" because of Morewedge et al, 2015; Khan et al, 2006; Baumer et al, 2015 [private slack conversation link](https://rethinkpriorities.slack.com/archives/G01962YABHB/p1655407493604669)
